maze-dataset: Maze Generation with Algorithmic Variety and Representational Flexibility

### 2025-04-09 docs/paper/paper.tex
fix some typos

\begin{table}[H]
\centering
\begin{tabular}{|ll|r|rrr|}
  \hline
  maze\_ctor
          & keyword args           & all sizes 
                                              & \shortstack{small \\ $g \leq 10$} 
                                                         & \shortstack{medium \\ $g \in (10, 32]$} 
                                                                    & \shortstack{large \\ $g > 32$} \\
  \hline\hline
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
          &                        &   28.0   &    2.8   &   20.3   &  131.8   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
          & accessible\_cells=20   &    2.3   &    2.2   &    2.4   &    2.2   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
          & do\_forks=False        &    2.7   &    2.2   &    3.1   &    3.5   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
          & max\_tree\_depth=0.5   &    2.5   &    2.0   &    2.7   &    4.0   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
          & p=0.1                  &   43.9   &    2.8   &   33.9   &  208.0   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
          & p=0.4                  &   48.7   &    3.0   &   36.5   &  233.5   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_kruskal}{kruskal}
          &                        &   12.8   &    1.9   &   10.3   &   55.8   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_percolation}{percolation}
          & p=1.0                  &   50.2   &    2.6   &   37.2   &  242.5   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_recursive_division}{recursive\_div}
          &                        &   10.2   &    1.7   &    8.9   &   42.1   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_wilson}{wilson}
          &                        &  676.5   &    7.8   &  188.6   & 3992.6   \\
  \hline\hline
  mean
          &                        &  559.9   &   13.0   &  223.5   & 3146.9   \\
  median
          &                        &   11.1   &    6.5   &   32.9   &  302.7   \\
  \hline
\end{tabular}
\caption{Generation times for various algorithms and maze sizes.}
\label{tab:benchmarks}
\end{table}

@@

The base function learned by symbolic regression privdes limited insight
and may be subject to change. It is defined as
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#cfg_success_predict_fn}{\texttt{cfg\_success\_predict\_fn}},
and takes a 5 dimensional float vector created by
\texttt{MazeDatasetConfig.\_to\_ps\_array()} which represents the 0)
percolation value 1) grid size 2) endpoint deadend configuration 3)
endpoint uniqueness 4) categorical generation function index.

@@

We combine these to get the
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
function, which is identity-like for \(p \approx 0.5\), and pushes
pushes \(x\) to extremes otherwise. \[
  \text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))
\]

@@

Instead, we describe mazes with the following representation: for a
\(2\)-dimensional lattice with \(r\) rows and \(c\) columns, we
initialize a boolean array \[
  A = \{0, 1\}^{2 \times r \times c}
\] which we refer to in the code as a
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
The value at \(A[0,i,j]\) determines whether a \emph{downward}
connection exists from node \([i,j]\) to \([i+1, j]\). Likewise, the
value at \(A[1,i,j]\) determines whether a \emph{rightward} connection
to \([i, j+1]\) exists. Thus, we avoid duplication of data about the
existence of connections and facillitate fast lookup time, at the cost
of requiring additional care with indexing. Note that this setup allows
for a periodic lattice. Generation of mazes is detailed in
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}.

>>>>

\begin{table}[H]
\centering
\begin{tabular}{|ll|r|rrr|}
  \hline
  maze\_ctor
          & keyword args           & all sizes 
                                              & \shortstack{small \\ $g \leq 10$} 
                                                         & \shortstack{medium \\ $g \in (10, 32]$} 
                                                                    & \shortstack{large \\ $g > 32$} \\
  \hline\hline
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
          &                        &   28.0   &    2.8   &   20.3   &  131.8   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
          & accessible\_cells=20   &    2.3   &    2.2   &    2.4   &    2.2   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
          & do\_forks=False        &    2.7   &    2.2   &    3.1   &    3.5   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs}{dfs}
          & max\_tree\_depth=0.5   &    2.5   &    2.0   &    2.7   &    4.0   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
          & p=0.1                  &   43.9   &    2.8   &   33.9   &  208.0   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_dfs_percolation}{dfs\_percolation}
          & p=0.4                  &   48.7   &    3.0   &   36.5   &  233.5   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_kruskal}{kruskal}
          &                        &   12.8   &    1.9   &   10.3   &   55.8   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_percolation}{percolation}
          & p=1.0                  &   50.2   &    2.6   &   37.2   &  242.5   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_recursive_division}{recursive\_div}
          &                        &   10.2   &    1.7   &    8.9   &   42.1   \\
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators.gen_wilson}{wilson}
          &                        &  676.5   &    7.8   &  188.6   & 3992.6   \\
  \hline\hline
  mean
          &                        &  559.9   &   13.0   &  223.5   & 3146.9   \\
  median
          &                        &   11.1   &    6.5   &   32.9   &  302.7   \\
  \hline
\end{tabular}
\caption{Generation times for various algorithms and maze sizes. More information can be found on the \docsref{benchmarks}{benchmarks page}.}
\label{tab:benchmarks}
\end{table}

@@

The base function learned by symbolic regression provides limited
insight and may be subject to change. It is defined as
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#cfg_success_predict_fn}{\texttt{cfg\_success\_predict\_fn}},
and takes a 5 dimensional float vector created by
\texttt{MazeDatasetConfig.\_to\_ps\_array()} which represents the
{[}percolation value, grid size, endpoint deadend configuration,
endpoint uniqueness, categorical generation function index{]}.

@@

We combine these to get the
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
function, which is identity-like for \(p \approx 0.5\), and pushes \(x\)
to extremes otherwise. \[
  \text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))
\]

@@

Instead, we describe mazes with the following representation: for a
\(2\)-dimensional lattice with \(r\) rows and \(c\) columns, we
initialize a boolean array \[
  A = \{0, 1\}^{2 \times r \times c}
\] which we refer to in the code as a
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
The value at \(A[0,i,j]\) determines whether a \emph{downward}
connection exists from node \([i,j]\) to \([i+1, j]\). Likewise, the
value at \(A[1,i,j]\) determines whether a \emph{rightward} connection
to \([i, j+1]\) exists. Thus, we avoid duplication of data about the
existence of connections and facilitate fast lookup time, at the cost of
requiring additional care with indexing. Note that this setup allows for
a periodic lattice. Generation of mazes is detailed in
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}.
